SPENCER THESIS - WORK IN PROGRESS

### 2016-05-28 doc/thesis_swe.tex
Thesis update.

%%------------------------------------------------------------------------------------------------------------------------------------------
%% FIGURE: Millennium Simulation Distribution plots (cosmic web, galaxy clusters with/without dark matter)
\begin{figure*}
    \centering
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_galaxies.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_galaxies}
    \end{subfigure}
    ~
    %% Einstein Ring Figure
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_halos.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_halos}
    \end{subfigure}
    \caption{}
\end{figure*}

@@

Cosmological simulations like the Millennium Simulation (MS) demonstrate that the underlying dark matter distribution of our universe is likely breathtakingly complex; rather than a largely disordered collection of galaxies and interstellar gas, matter appears to condense into clusters and superclusters of galaxies threaded together by large filaments (pocketed) with under-dense regions called voids, as shown in Figure ??(1.a) (CITE MS). While the exact relation between the distribution of galaxies and dark matter is not known, simulations predict that early galaxies formed in the large gravitational potentials of over-dense regions of dark matter (CITE!). This suggests that, while we cannot map dark matter directly, galaxies should trace out at least some of the underlying dark matter structures \textcolor{red}{(Citation!)}. A comparrison of the predicted galaxy and dark matter distribution from a small region of the MS is shown in Figure (???).

@@

The simplest way to model the relationship between galaxies and the dark matter distribution is by enveloping each galaxy in a spherically symmetric dark matter `halo' of a certain mass $M_h$. These halos extend far beyond the edge of the visible galaxy that they enclose, with the Milky Way's own halo estimated to have a radius of \textcolor{red}{(Check and cite!)}. While the density profile of the halos may be complex, numerous simulations have shown that it can be well approximated locally by the Navarro-Frenk-White (NFW) profile which has the form
\begin{equation}\label{nfw_profile}
\rho_{NFW}(r)=\frac{\rho_0}{\frac{r}{R_s}\left(1+\frac{r}{R_s}\right)^2}
\end{equation}

\noindent \textcolor{red}{where the constant $\rho_0$ and the scale radius $R_s$ are parameters that vary from halo to halo} \cite{nfw}. However, the NFW profile is not entierly physical as the total mass diverges as the desntiy is integrated out to an infinite radius, so this work uses a truncated NFW profile called the Baltz-Marshall-Oguri (BMO) profile given by
\begin{equation}\label{bmo_profile}
\rho_{BMO}(r)=\rho_{NFW}\cdot\left(\frac{r_t^2}{r^2+r_t^2}\right)^2
\end{equation}

@@

The halo model only accounts for dark matter structures associated with galaxies. However, this neglects other structures such as the filaments and voids that can be seen in Figure (??). Modelling these features accurately would be difficult; we could try, for example, to model filaments by attaching cylinders of dark matter between massive clusters of galaxies. While an intriguing idea, rigorously testing for the appropriate filament density distributions, selection criteria, and frequency to statistically match simulations could constitute its own thesis.

@@

Instead, as the mean mass density of large enough ($\sim$100 Mpc) regions should approximate the mean mass density of the universe, we can make a `first-order' approximation by assuming that all additional mass (or absence of mass) is contained in a uniform `smooth-component' density that ensures the total mass density of any region equals the mean density of the universe at that redshift. This assumption can be expressed as
\begin{equation}\label{smooth_correction1}
\rho_{\text{matter}}(z)=\rho_{\text{halos}}(z)+\rho_{\text{smooth}}(z)
\end{equation}

@@

Using the fluid equation and equation of state \textcolor{red}{(CITE)}, we can show that the energy density of non-relativistic matter $\varepsilon_m(z)$ evolves as
\begin{equation}\label{energy_evol}
\varepsilon_m(z)=\frac{\varepsilon_{m,0}}{a(z)^3}=\varepsilon_{m,0}(1+z)^3
\end{equation}

@@

The halo mass density $\rho_h(z)$ at a given redshift is simply the sum of the individual halo masses $M_{h,i}$ over a physical volume $dV_p$. However, we want a proper volume element that is expressed in terms of the solid angle $d\Omega$ and the redshift $dz$. To do this, we can start with the standard spherical volume element $dV=r^2drd\Omega$. However, we must be careful that (1) we are using proper distances, (2) account for the possibility of a curved universe, and (3) use the correct distance measure for $r$ (in this case, the angular diameter distance). From this we can infer that the proper volume element must have the form of
\begin{equation}\label{proper_element}
dV_p=d_A(z)^2dr_pd\Omega
\end{equation}

@@

While dark matter does not absorb or emit light, it does interact with light indirectly through gravitational lensing. The presence of matter curves spacetime locally which causes light to follow curved paths called geodesics \textcolor{red}{(clean up!)}. The deviation in path is usually negligable, but can be significant in regions containing clusters of galaxies and their massive dark matter halos. By observing how the magnitudes and shapes of distant galaxies are distorted when travelling through regions of matter, we can infer the masses of the dark matter halos and reconstruct the hidden dark matter mass distribution. A simplified cartoon of this process is shown in Figure ??, as well as a diagram defining the necessary angles and distances in Figure ??.

@@

\noindent where $\sum(\vec{r},z_l)$ is the projected surface mass density of the lensing mass and $\sum_{cr}(z_l,z_s)$ is the critical surface density for a given source redshift $z_s$ and lens redshift $z_l$ (\textcolor{red}{Citation}). The shearing of the source is most often described by the complex shear $\gamma$ defined to be
\begin{equation}\label{complex_shear}
\gamma=\gamma_1+i\gamma_2=|\gamma|\,e^{2i\varphi}
\end{equation}

@@

(introduce issue)

@@

\noindent and $r\leq1$ is the ratio of the semi-minor and semi-major axes of the ellipse. This compact notation combines the eccentricity and orientation of the ellipse into a single object. A plot from Schneider et al. \cite{schneider} showing the shape of elliptical galaxies for various values of $\varepsilon_1$ and $\varepsilon_2$ is shown in Figure \ref{ellipses}.

There are many complications to using this scheme in practice, most notably the multiplicative bias resulting from the smearing of galaxy images by the observational point spread function (PSF) \cite{multiplicative_bias}. While the effects of a PSF can be complex, in general it causes galaxy images to appear less elliptical than they truly are. To account for this in generated galaxy images, a multiplicative bias parameter $M$ is often used to lessen the intrinsic ellipticity using
$$\varepsilon_{obs}=M\cdot\varepsilon_{int}$$

\noindent where $\varepsilon_{int}$ is the generated intrinsic ellipticity of the image and $\varepsilon_{obs}$ is the ellipticity that would be recorded by a detector.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Cartoon
\begin{figure*}
    \centering
    \includegraphics[width=0.85\textwidth]{figs-swe/thesis/pangloss_cartoon.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{A cartoon model of how the \texttt{Pangloss} framework uses a dark matter halo model to reconstruct the foreground mass along the line of sight of a background source galaxy. A `lightcone' centered at a background source is constructed with radius $R$ and populated with all foreground galaxies contained in this volume. Each foreground object has an attached dark matter halo and physical distance to the line of sight (dashed line). The shear and convergence along the line of sight contributed by each halo is calculated using \cite{lensing_calc}, and the predicted ellipticity is determined using the sum of these contributions and Equation \eqref{lensing}.}
    \label{pangloss_cartoon}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

To infer foreground dark matter mass structures using weak gravitational lensing, first a model of the relationship between foreground galaxies and the foreground dark matter must be established and robustly tested to see if, statistically, it makes the same lensing predictions of background sources as the true underlying dark matter structure. To do this, I built upon the publicly available \texttt{Pangloss}\footnote{\label{note1}https://github.com/drphilmarshall/Pangloss} framework used in Collett et al. \cite{collett_marshall} to reconstruct all the
mass along the line of sight of each background galaxy using a halo model with smooth-component correction. The lensing contribution of each halo is calculated, and the total lensing of the background galaxy is the sum of each halo contribution. The implementation of this process is detailed in the following sections.

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Model Assumptions} \label{assumptions}
While \texttt{Pangloss} may be used more generally, the present analysis makes some additional strong assumptions to simplify the problem for a first attempt at making weak lensing predictions.

@@

Testing the first assumption is the main goal of this paper. The second assumption is reasonable as it is estimated that dark matter halos are on average 1-2 orders of magnitude more massive than the host galaxy \cite{smhr}. Clearly the third assumption will never be true for any observational data. However, this allows for a best-case scenario estimate of how well the \texttt{Pangloss} framework \textit{could} predict weak lensing effects given all possible information. This assumption can be relaxed by sampling a dark matter halo mass from an assumed stellar mass - halo mass relation. The fourth assumption is also unrealistic as most galaxies in sky surveys only have a less-reliable photometric redshift due to time constraints, but this again allows for a best-case estimate. This assumption could be relaxed by instead using photometric redshifts, adding random noise, and repeating the upcoming analysis on many realizations of the photometric redshifts.

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{The Millennium Simulation} \label{ms}
\texttt{Pangloss} cannot be used to make dark matter mass maps using existing galaxy catalogs until it is tested on a simulated universe with known dark matter structure to determine how accurately and precisely it predicts the lensing of background sources. For this purpose, \texttt{Pangloss} was tested on galaxy catalogs from the Millennium Simulation, an N-body simulation consisting of over 10 billion dark matter `particles' each representing a billion solar masses and populated with about 20 million galaxies \cite{millennium_simulation}. The simulation contains baryonic and dark matter structure believed to be similar to our own universe, and uses cosmological parameters from WMAP 1\textsuperscript{st}-year data analysis which assumed a matter density of $\Omega_m=0.25$, a dark energy density of $\Omega_\Lambda=0.75$, and a Hubble constant of $H_0=73$ km$\,$s$^{-1}\,$Mpc$^{-1}$ (\textcolor{red}{CITE!!}). From the work of Hilbert et al. \cite{ray_tracing}, high resolution maps of the ray-traced shear and convergence of galaxy sources at $z=1.3857$ are publicly available. From these maps, an estimate of the `true' lensing of background galaxy ellipticities when traveling through the foreground mass of the Millennium Simulation can be calculated using Equation \eqref{lensing}.

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Generating Background Galaxies}
With a catalog of foreground galaxies chosen, a set of \textcolor{red}{???} background galaxies over a field of \textcolor{red}{???} arcmin$^2$ was generated, with density 10 per arcmin$^2$. The intrinsic orientation of each galaxy was sampled from a uniform distribution as, without lensing, there should be no preferred orientation due to the assumption of an isotropic universe. The magnitude of the galaxy ellipticities was sampled from a normal distribution with a standard deviation of 0.2, but any ellipticities with magnitude greater than one were re-sampled. Random ellipticity noise was sampled from a normal distribution with a standard deviation of 0.1 and added to the intrinsic ellipticities. Finally, each ellipticity was multiplied by $M=0.9$ to account for a multiplicative bias of 10\%.

@@

While ideally all foreground mass in a field would be considered when predicting the weak lensing of a background galaxy, it is computationally prohibitive to do so. Instead, all foreground halos contained within a `lightcone' centered along the line of sight to the source and extending out to a chosen lightcone radius $R$ were considered when calculating the lensing contributions for the background source. A cartoon model of this process is shown in Figure \ref{pangloss_cartoon}. Unless otherwise specified, experiments in this paper used a radius of 6 arcminutes and halo truncation scale of five times the virial radius.

To calculate the convergence and shear contribution of each halo, the physical distance from the halo to the line of sight was needed. To increase the speed of distance calculations, the foreground halo redshifts were first binned to a grid of 100 equally-sized redshift bins $\Delta z\approx0.013857$ from $z=0$ to the sourced redshift of $z=1.3857$, and then converted to physical distances using the cosmology defined by the Millennium Simulation.

@@

Not all foreground halos are equally relevant to the lensing calculation. To add a significant contribution to the combined shear and convergence, a halo must be massive, close to the line of sight, or preferably both. McCully et al. \textcolor{red}{(CITE)} derived that the correct metric for determining the relevance of a given foreground halo to the overall lensing calculation to be given by
\begin{equation}\label{relevant}
Rel(M_h,R)\propto\frac{M_h}{R^3}
\end{equation}

@@

In this work, the threshold values used are $M_T=10^{12}M_\odot$ and $R_T=10$ kpc. The relevance distribution for one background catalog realization with \textcolor{red}{???} foreground halos in a lightcone of radius \textcolor{red}{??} arcminutes is shown in Figure ??. As expected, there are very few massive halos along the line of sight and therefore only a small percentage of foreground halos should have a high relative contribution to the predicted lensing quantities. This encouraged the intended analysis on lightcones with all halos and those with only a small subset of the most relevant halos to compare the results and see if the computational time saved by using only the most relevant halos was worth the loss in prediction accuracy.

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Relevance Distribution of all Halos
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/relevance_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:rel_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

As discussed in Section \ref{scc}, the halo model does not account for additional mass structures such as filaments. To apply the smooth-component correction to \texttt{Pangloss}, the mean mass density $\rho(z)$ and halo density $\rho_h(z)$ had to be calculated at each of the 100 redshift bins. For $\rho_h(z)$, this was done by summing all halo masses $M_{h,i}$ in a given redshift bin and integrating Equation \eqref{pvolume_element} from ${z-\Delta z/2}$ to ${z+\Delta z/2}$, except for the boundary bins. Once $\rho_h(z)$ had been calculated for each redshift slice, the projected surface mass density was found by integrating over the redshift interval,
\begin{align}\label{surface_density} 
\sum(z)&=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho_s(z)\,dz\nonumber\\
&=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho(z)-\rho_h(z)\,dz,
\end{align}

@@

The volume and projected surface mass densities of each redshift slice are shown in Figures \ref{fig:volume_densities} and \ref{fig:surface_densities} respectively. (Describe plots) (reduced shears before and after correction)

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
\begin{figure*}
    \centering
    % Volume
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/vol_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:volume_densities}
    \end{subfigure}
    ~
    % Surface
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/surf_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:surface_densities}
    \end{subfigure}
    \caption{}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Kappa Contributions of S.C.C.
\begin{figure}
    \centering
    \includegraphics[width=0.425\textwidth]{figs-swe/thesis/kappa_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:kappa_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%==========================================================================================================================================
\section{Checking the Pangloss Model}

@@

The catalog of lensed ellipticities was also analyzed using the galaxy-mass correlation function with separation distances from ?? to ?? arcminutes and a lightcone radius of ?? arcminutes. The result is shown in Figure \ref{ng_corr}. As with the ellipticity-ellipticity correlation function, there is no significant correlation for the intrinsic ellipticities at any scale as expected. The lensed ellipticities from the ray-tracing and halo model both show positive correlation on all calculated scales and are in agreement on separations from ?? to ?? arcminutes. However, .... (New analysis)

%the halo model is still systematically underpredicting the shear power on scales above 0.7 arcminutes, although now only by an average of 23\%. A new feature is the large overprediction of shear power on separation scales smaller than 0.2 arcminutes. However, background sources this close to foreground halos are often in the strong lensing regime which is not currently accounted for by the \texttt{Pangloss} framework.

The galaxy-mass correlation function was similarly re-calculated at increasing lightcone radii from $R=1$ to $R=8$ arcminutes with the result in Figure ??. (Analyze new figure...)

>>>>

%%------------------------------------------------------------------------------------------------------------------------------------------
%% FIGURE: Millennium Simulation Distribution plots (cosmic web, galaxy clusters with/without dark matter)
\begin{figure*}
    \centering
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_galaxies.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_galaxies}
    \end{subfigure}
    ~
    %% Einstein Ring Figure
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_halos.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_halos}
    \end{subfigure}
    \caption{}
    \label{ms_compare}
\end{figure*}

@@

Cosmological simulations like the Millennium Simulation (MS) demonstrate that the underlying dark matter distribution of our universe is likely breathtakingly complex; rather than a largely disordered collection of galaxies and interstellar gas, matter appears to condense into clusters and superclusters of galaxies threaded together by large filaments (pocketed) with under-dense regions called voids, as shown in Figure ??(1.a) (CITE MS). While the exact relation between the distribution of galaxies and dark matter is not known, simulations predict that early galaxies formed in the large gravitational potentials of over-dense regions of dark matter (CITE!). This suggests that, while we cannot map dark matter directly, galaxies should trace out at least some of the underlying dark matter structures (Citation!). A comparrison of the predicted galaxy and dark matter distribution from a small region of the MS is shown in Figure \ref{ms_compare}.

@@

The simplest way to model the relationship between galaxies and the dark matter distribution is by enveloping each galaxy in a spherically symmetric dark matter `halo' of a certain mass $M_h$. These halos extend far beyond the edge of the visible galaxy that they enclose, with the Milky Way's own halo estimated to have a radius of (Check and cite!). While the density profile of the halos may be complex, numerous simulations have shown that it can be well approximated locally by the Navarro-Frenk-White (NFW) profile which has the form
\begin{equation}\label{nfw_profile}
\rho_{NFW}(r)=\frac{\rho_0}{\frac{r}{R_s}\left(1+\frac{r}{R_s}\right)^2}
\end{equation}

\noindent where the constant $\rho_0$ and the scale radius $R_s$ are parameters that vary from halo to halo \cite{nfw}. However, the NFW profile is not entierly physical as the total mass diverges as the desntiy is integrated out to an infinite radius, so this work uses a truncated NFW profile called the Baltz-Marshall-Oguri (BMO) profile given by
\begin{equation}\label{bmo_profile}
\rho_{BMO}(r)=\rho_{NFW}\cdot\left(\frac{r_t^2}{r^2+r_t^2}\right)^2
\end{equation}

@@

The halo model only accounts for dark matter structures associated with galaxies. However, this neglects other mass structures such as the filaments and voids that can be seen in Figure (??). Modelling these features accurately would be difficult; we could try, for example, to model filaments by attaching cylinders of dark matter between massive clusters of galaxies. While an intriguing idea, rigorously testing for the appropriate filament density distributions, selection criteria, and frequency to statistically match simulations could constitute its own thesis.

@@

Instead, as the mean mass density of large enough ($\sim$100 Mpc) regions should approximate the mean mass density of the universe, we can make a `first-order' approximation by assuming that all additional mass (or absence of mass) is contained in a uniform `smooth-component' density that ensures the total mass density of any sufficiently large region equals the mean density of the universe at that redshift. This assumption can be expressed as
\begin{equation}\label{smooth_correction1}
\rho_{\text{matter}}(z)=\rho_{\text{halos}}(z)+\rho_{\text{smooth}}(z)
\end{equation}

@@

Using the fluid equation and equation of state (Cite), we can show that the energy density of non-relativistic matter $\varepsilon_m(z)$ evolves as
\begin{equation}\label{energy_evol}
\varepsilon_m(z)=\frac{\varepsilon_{m,0}}{a(z)^3}=\varepsilon_{m,0}(1+z)^3
\end{equation}

@@

The halo mass density $\rho_h(z)$ at a given redshift is simply the sum of the individual halo masses $M_{h,i}$ over a physical volume $dV_p$. However, we want a proper volume element that is expressed in terms of the solid angle $d\Omega$ and the redshift $dz$. To do this, we can start with the standard spherical volume element $dV=r^2drd\Omega$. However, one must be careful to (1) use proper distances, (2) account for the possibility of a curved universe, and (3) use the correct distance measure for $r$ (in this case, the angular diameter distance). From this one can infer that the proper volume element must have the form of
\begin{equation}\label{proper_element}
dV_p=d_A(z)^2dr_pd\Omega
\end{equation}

@@

While dark matter does not absorb or emit light, it does interact with light indirectly through gravitational lensing. The presence of matter curves spacetime locally which causes light to follow curved paths called geodesics. The deviation in path is usually negligable, but can be significant in regions containing clusters of galaxies and their massive dark matter halos. By observing how the magnitudes and shapes of distant galaxies are distorted when travelling through regions of matter, we can infer the masses of the dark matter halos and reconstruct the hidden dark matter mass distribution. A simplified cartoon of this process is shown in Figure ??, as well as a diagram defining the necessary angles and distances in Figure ??.

@@

\noindent where $\sum(\vec{r},z_l)$ is the projected surface mass density of the lensing mass and $\sum_{cr}(z_l,z_s)$ is the critical surface density for a given source redshift $z_s$ and lens redshift $z_l$ (Citation). The shearing of the source is most often described by the complex shear $\gamma$ defined to be
\begin{equation}\label{complex_shear}
\gamma=\gamma_1+i\gamma_2=|\gamma|\,e^{2i\varphi}
\end{equation}

@@

To calculate the lensed ellipticities in the previous section, galaxies must first be translated from pixel intensities into elliptical representations of the galaxies. This is not always possible for galaxies are diverse in type and shape and not all are well approximated by ellipses. However for this thesis, all galaxy ellipticities are simulated so this issue is avoided. For a more thorough analysis of measuring image ellipticities, see Schneider (2003) \cite{schneider}.

@@

\noindent and $r\leq1$ is the ratio of the semi-minor and semi-major axes of the ellipse. This compact notation combines the eccentricity and orientation of the ellipse into a single object. A plot from Schneider (2003) \cite{schneider} showing the shape of elliptical galaxies for various values of $\varepsilon_1$ and $\varepsilon_2$ is shown in Figure \ref{ellipses}.

There are many complications to using this scheme in practice, most notably the multiplicative bias resulting from the smearing of galaxy images by the observational point spread function (PSF) \cite{multiplicative_bias}. While the effects of a PSF can be complex, in general it causes galaxy images to appear less elliptical than they truly are. To account for this in generated galaxy images, a multiplicative bias parameter $M$ is often used to lessen the intrinsic ellipticity using
\begin{equation}\label{mult_bias}
\varepsilon_{\text{obs}}=M\cdot\varepsilon_{\text{int}}
\end{equation}

\noindent where $\varepsilon_{\text{int}}$ is the generated intrinsic ellipticity of the image and $\varepsilon_{\text{obs}}$ is the ellipticity that would be recorded by a detector.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Cartoon
\begin{figure*}
    \centering
    \includegraphics[width=0.85\textwidth]{figs-swe/thesis/pangloss_cartoon.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{A cartoon model of how the \texttt{Pangloss} framework uses a dark matter halo model to reconstruct the foreground mass along the line of sight of a background source galaxy. A `lightcone' centered at a background source is constructed with radius $R_L$ and populated with all foreground galaxies contained in this volume. Each foreground object has an attached dark matter halo and physical distance to the line of sight (dashed line). The shear and convergence along the line of sight contributed by each halo is calculated using \cite{lensing_calc}, and the predicted ellipticity is determined using the sum of these contributions and Equation \eqref{lensing}.}
    \label{pangloss_cartoon}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

To infer foreground dark matter mass structures using weak gravitational lensing, first a model of the relationship between foreground galaxies and the foreground dark matter must be established and robustly tested to see if, statistically, it makes the same lensing predictions of background sources as the true underlying dark matter structure. To do this, I built upon the publicly available \texttt{Pangloss}\footnote{\label{note1}https://github.com/drphilmarshall/Pangloss} framework used in Collett et al. \cite{collett_marshall} to reconstruct all the
mass along the lines of sight of generated background galaxies in the MS using a halo model with smooth-component correction. The lensing contribution of each halo is calculated, and the total lensing of the background galaxy is the sum of each halo contribution. The implementation of this process is detailed in the following sections.

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Model Assumptions} \label{assumptions}
While \texttt{Pangloss} may be used more generally, the present analysis makes some additional strong assumptions to simplify the problem for a first attempt at making weak lensing predictions:

@@

Testing the first assumption is the main goal of this paper. The second assumption is reasonable as it is estimated that dark matter halos are on average 1-2 orders of magnitude more massive than the host galaxy \cite{smhr}. Clearly the third assumption will never be true for any observational data. However, this allows for a best-case estimate of how well the \texttt{Pangloss} framework \textit{could} predict weak lensing effects given all possible information. This assumption can be relaxed by sampling a dark matter halo mass from an assumed stellar mass - halo mass relation. The fourth assumption is also unrealistic as most galaxies in sky surveys only have a less-reliable photometric redshift due to time constraints, but this again allows for a best-case estimate. This assumption could be relaxed by instead using photometric redshifts, adding random noise, and repeating the upcoming analysis on many realizations of the photometric redshifts.

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{The Millennium Simulation} \label{ms}
\texttt{Pangloss} cannot be used to make dark matter mass maps using existing galaxy catalogs until it is tested on a simulated universe with known dark matter structure to determine how accurately and precisely it predicts the lensing of background sources. For this purpose, \texttt{Pangloss} was tested on galaxy catalogs from the Millennium Simulation, an N-body simulation consisting of over 10 billion dark matter `particles' each representing a billion solar masses and populated with about 20 million galaxies \cite{millennium_simulation}. The simulation contains baryonic and dark matter structure believed to be similar to our own universe, and uses cosmological parameters from WMAP 1\textsuperscript{st}-year data analysis which assumed a matter density of $\Omega_m=0.25$, a dark energy density of $\Omega_\Lambda=0.75$, and a Hubble constant of $H_0=73$ km$\,$s$^{-1}\,$Mpc$^{-1}$ (Cite). From the work of Hilbert et al. \cite{ray_tracing}, high resolution maps of the ray-traced shear and convergence of galaxy sources at $z=1.3857$ are publicly available. From these maps, an estimate of the `true' lensing of background galaxy ellipticities when traveling through the foreground mass of the Millennium Simulation can be calculated using Equation \eqref{lensing}.

%%------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Generating Background Galaxies}
With the (0,0,0,0) catalog of MS foreground galaxies chosen, a set of ?? background galaxies over a field of ?? arcmin$^2$ was generated, with density 10 per arcmin$^2$. The intrinsic orientation of each galaxy was sampled from a uniform distribution as, without lensing, there should be no preferred orientation due to the assumption of an isotropic universe. The magnitude of the galaxy ellipticities was sampled from a normal distribution with a standard deviation of 0.2, but any ellipticities with magnitude greater than one were re-sampled. Random ellipticity noise was sampled from a normal distribution with a standard deviation of 0.1 and added to the intrinsic ellipticities. Finally, each ellipticity was multiplied by $M=0.9$ to account for a multiplicative bias of 10\%.

@@

While ideally all foreground mass in a field would be considered when predicting the weak lensing of a background galaxy, it is computationally prohibitive to do so. Instead, all foreground halos contained within a `lightcone' centered along the line of sight to the source and extending out to a chosen lightcone radius $R_L$ were considered when calculating the lensing contributions for the background source. A cartoon model of this process is shown in Figure \ref{pangloss_cartoon}. Unless otherwise specified, experiments in this paper used a radius of 6 arcminutes and halo truncation scale of five times the virial radius.

To calculate the convergence and shear contribution of each halo, the physical distance from the halo to the line of sight was needed. To increase the speed of distance calculations, the foreground halo redshifts were first binned to a grid of 100 equally-sized redshift bins $\Delta z=0.013857$ from $z=0$ to the sourced redshift of $z=1.3857$, and then converted to physical distances using the cosmology defined by the Millennium Simulation.

@@

Not all foreground halos are equally relevant to the lensing calculation. To add a significant contribution to the combined shear and convergence, a halo must be massive, close to the line of sight, or preferably both. McCully et al. (Cite) derived that the correct metric for determining the relevance of a given foreground halo to the overall lensing calculation to be given by
\begin{equation}\label{relevant}
Rel(M_h,R)\propto\frac{M_h}{R^3}
\end{equation}

@@

In this work, the threshold values used are $M_T=10^{12}M_\odot$ and $R_T=10$ kpc. The relevance distribution for one background catalog realization with ?? foreground halos in a lightcone of radius ?? arcminutes is shown in Figure ??. As expected, there are very few massive halos along the line of sight and therefore only a small percentage of foreground halos should have a high relative contribution to the predicted lensing quantities. This encouraged the intended analysis on lightcones with all halos and those with only a small subset of the most relevant halos to compare the results and see if the computational time saved by using only the most relevant halos was worth the loss in prediction accuracy.

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Relevance Distribution of all Halos
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs-swe/thesis/relevance_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:rel_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

As discussed in Section \ref{scc}, the halo model does not account for additional mass structures such as filaments. To apply the smooth-component correction to \texttt{Pangloss}, the mean mass density $\rho(z)$ and halo density $\rho_h(z)$ had to be calculated at each of the 100 redshift bins. For $\rho_h(z)$, this was done by summing all halo masses $M_{h,i}$ in a given redshift bin and integrating Equation \eqref{pvolume_element} from ${z-\Delta z/2}$ to ${z+\Delta z/2}$, except for the boundary bins, and multiplying by the solid angle of a cone ${\Omega=2\pi(1-\cos R_L)}$. Once $\rho_h(z)$ had been calculated for each redshift slice, the projected surface mass density was found by integrating over the redshift interval,
\begin{align}\label{surface_density} 
\sum(z)&=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho_s(z)\,dz\nonumber\\
&=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho(z)-\rho_h(z)\,dz,
\end{align}

@@

The volume and projected surface mass densities of each redshift slice are shown in Figures \ref{fig:volume_densities} and \ref{fig:surface_densities} respectively. In each plot, the halo density at each redshift slice is plotted in green, the smooth-component density in blue, and universe mean density in black, and a histogram of the galaxies per bin is shown in the background. In most redshift bins, the halo density is one to two orders of magnitude lower than than the universe mean density leading to a smooth-component density very close to that of the mean density. This seems to be in opposition to Pache et al. (year) \cite{pache} who found that halos constituted ??\% of simulation mass in ??. However, all mass densities are far below the critial density, plotted in red, and so well within the weak-lensing regime. The convergence distribution of each halo and 100 smooth-components is plotted in Figure \ref{fig:kappa_dist}.
%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
\begin{figure*}
    \centering
    % Volume
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/vol_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:volume_densities}
    \end{subfigure}
    ~
    % Surface
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/surf_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:surface_densities}
    \end{subfigure}
    \caption{}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Kappa Contributions of S.C.C.
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/kappa_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:kappa_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%==========================================================================================================================================
\section{Checking the Pangloss Model} \label{check}

@@

The catalog of lensed ellipticities was also analyzed using the galaxy-mass correlation function with separation distances from ?? to ?? arcminutes and a lightcone radius of ?? arcminutes. The result is shown in Figure \ref{ng_corr}. The lensed ellipticities from the ray-tracing and halo model both show positive correlation on all calculated scales and are in agreement on separations from ?? to ?? arcminutes. However, the halo model is still systematically underpredicting the shear power on scales above ?? arcminutes, although now only by an average of ??\%. A new feature is the large overprediction of shear power on separation scales smaller than ?? arcminutes. This is likely because background sources this close to foreground halos are often in the strong lensing regime which is not currently accounted for by the \texttt{Pangloss} framework. The NRMSE of the halo correlation is ??.

The galaxy-mass correlation function was similarly re-calculated at increasing lightcone radii from $R=1$ to $R=8$ arcminutes with the result in Figure ??. As with the $\varepsilon$-$\varepsilon$ correlation, ...
